\chapter*{Foreword}

\vk
\paragraph{Welcome!} This manual describes how to use the Laboratoire de M\'et\'eorologie Dynamique (LMD) Martian Mesoscale Model. Many thanks for looking forward to using this model which development required countless hours of hard work! A significant part of the model development and validation have been funded by ESA and CNES which are acknowledged for their support.

\paragraph{Contact} The main contact to reach at LMD to become an user of the model is Aymeric SPIGA (main developper, \href{mailto:aymeric.spiga@upmc.fr}{\nolinkurl{aymeric.spiga@upmc.fr}}). Alternative contacts at LMD for mesoscale modeling inquiries are Ehouarn MILLOUR~\url{ehouarn.millour@lmd.jussieu.fr} or Fran\c cois FORGET~\url{francois.forget@lmd.jussieu.fr}. We are open to questions and suggestions on new scientific collaborations, teaching/outreach actions or contractual proposals.

\paragraph{Copyright (LMD)} The LMD Martian Mesoscale Model sources are made available on the condition that we make no representations or warranties regarding the reliability or validity of the model predictions nor the use to which such model predictions should be put, disclaim any and all responsibility for any errors or inaccuracies in the model predictions and bear no responsibility for any use made of this model predictions by any party. The scientific use of already published LMD Martian Mesoscale Model simulations is freely allowed provided that the reference paper \textit{Spiga and Forget} [2009]\nocite{Spig:09} is correctly quoted in all publications and that we are kept informed of usage and developments. If your study requires dedicated simulations with the LMD Martian Mesoscale Model, please consider including Aymeric SPIGA as a co-author of your work and asking, if needed, for help with writing the part related to mesoscale modeling. If your study requires additional work on a specific Martian physical parameterization, please consider including other members of the LMD team in addition to Aymeric SPIGA. The LMD Martian Mesoscale Model may not be put to any commercial use without specific authorization. 

\paragraph{Copyright (WRF)} Part of the LMD Martian Mesoscale Model is based on the terrestrial model WRF which is in the public domain. If you are an user of the LMD Martian Mesoscale Model, you are therefore an user of the WRF model. Please take a minute to fill in the WRF registration form so that the WRF development team knows about the people using their model: \url{http://www.mmm.ucar.edu/wrf/users/download/wrf-regist.php}. \noindent \scriptsize \emph{WRF was developed at the National Center for Atmospheric Research (NCAR) which is operated by the University Corporation for Atmospheric Research (UCAR). NCAR and UCAR make no proprietary claims, either statutory or otherwise, to this version and release of WRF and consider WRF to be in the public domain for use by any person or entity for any purpose without any fee or charge. UCAR requests that any WRF user include this notice on any partial or full copies of WRF. WRF is provided on an "AS IS" basis and any warranties, either express or implied, including but not limited to implied warranties of non-infringement, originality, merchantability and fitness for a particular purpose, are disclaimed. In no event shall UCAR be liable for any damages, whatsoever, whether direct, indirect, consequential or special, that arise out of or in connection with the access, use or performance of WRF, including infringement actions. WRF is a registered trademark of the University Corporation for Atmospheric Research (UCAR).} \normalsize

\clearemptydoublepage
